---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
activityData <- read.csv("./activity/activity.csv")
head(activityData)
```

## What is mean total number of steps taken per day?


```{r}
meanTotalPerDay <- aggregate(steps ~ date,activityData,mean,na.rm=T)
meanTotalPerDay
mean(meanTotalPerDay$steps)
```


## What is the average daily activity pattern?
Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and 
the average number of steps taken, averaged across all days (y-axis)
(Only complete cases, sind this data rame is used later to replace NAs in the source data frame)

```{r}
stepsPerInterval <- aggregate(steps ~ interval,activityData[(complete.cases(activityData)),],mean)
with(stepsPerInterval,plot(interval,steps,type="l"))
```

Which 5-minute interval, on average across all the days in the dataset, contains 
the maximum number of steps?

```{r}
maxSteps <- max(stepsPerInterval$steps)
maxInterval <- stepsPerInterval[stepsPerInterval$steps %in% maxSteps,"interval"]
maxInterval
```

## Imputing missing values
Calculate and report the total number of missing values in the dataset (i.e. 
the total number of rows with NAs)
```{r}
sum(!complete.cases(activityData))
```

Devise a strategy for filling in all of the missing values in the dataset. 
The strategy does not need to be sophisticated. For example, you could use 
the mean/median for that day, or the mean for that 5-minute interval, etc.
Create a new dataset that is equal to the original dataset but with the missing 
data filled in. 

My strategy: replace NAs by mean of steps per interval

```{r}
replaceNaByIntervalMean <- function(activityData) {
        for (i in 1:nrow(activityData)) {
                if (is.na(activityData[i,"steps"])) {
                        activityData[i,"steps"] <- stepsPerInterval[stepsPerInterval$interval == activityData[i,"interval"],"steps"] 
                } 
        }
        activityData
}

activityDataCC <- replaceNaByIntervalMean(activityData)
head(activityDataCC)
```

Make a histogram of the total number of steps taken each day and Calculate and 
report the mean and median total number of steps taken per day. 

```{r}
stepsPerDay <- aggregate(steps ~ date,activityData,sum)
stepsPerDayCC <- aggregate(steps ~ date,activityDataCC,sum)
hist(stepsPerDayCC$steps,breaks=50)
summary(stepsPerDayCC$steps)
hist(stepsPerDay$steps,breaks=50)
summary(stepsPerDay$steps)
```

Do these values differ from the estimates from the first part of the assignment? 
What is the impact of imputing missing data on the estimates of the total daily number of steps?
Yes, slightly. With replacing the NAs the quartile range is narrowed.

## Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable in the dataset with two levels -- "weekday" and 
"weekend" indicating whether a given date is a weekday or weekend day.

```{r}
weekdaysOfDates <- weekdays(as.Date(activityDataCC$date))
partOfWeek <- function(days) {
        days[days == "Samstag" | days == "Sonntag"] <- "weekend"
        days[days != "weekend"] <- "weekday"
        days
}
activityDataCC$partOfWeek <- as.factor(partOfWeek(weekdaysOfDates))
head(activityDataCC)
str(activityDataCC)
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 
5-minute interval (x-axis) and the average number of steps taken, averaged 
across all weekday days or weekend days (y-axis).

```{r}
stepsPerIntervalCC <- aggregate(steps ~ interval + partOfWeek,activityDataCC,mean)
library("ggplot2")
ggplot(stepsPerIntervalCC,aes(x=interval,y=steps)) + geom_line() + facet_grid(partOfWeek~.)
```
